---
title: "Algorithmic Trading"
author: "Alexander Martinez"
date: "6/13/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Loading the packages
```{r}
library(quantmod)
library(quantstrat)
library(data.table)
library(DT)
library(ggplot2)
library(htmltools)
library(htmlwidgets)
library(knitr)
library(lattice)
library(pander)
library(tidyverse)
sessionInfo()
```
#Setting system up
```{r}
# Set the time zone to UTC
Sys.setenv(TZ = 'UTC')
#Set currency to USD
currency("USD")
```
#Creating the Objects to Get started
I am going to use a time period that the market is weak. This is because we want to backtest to when times are bad as well as good to get current results.

init_date = date before start date of data retrieval
start_date = date to start data retrieval
end_date = date to end data retrieval 
init_equity = initial account equity
adjustment = TRUE if we should adjust prices for a various amiunt of factors
```{r}
init_date <- "2007-12-13"
start_date <- "2008-01-08"
end_date <- "2009-12-31"
init_equity <- 1e4 #10,000
adjustment <- TRUE
```
# Symbols
Most of the strategies will be focused on 3 ETFs: IWM, QQQ, SPY
```{r}
basic_symbols <- function(){
  symbols <- c("IWM", "QQQ", "SPY")
}
```
To test strategies on a broader scale, we will use enhanced_symbols()
```{r}
enhanced_symbols <- function(){
  symbols <- c(
    basic_symbols(),
    "TLT", # iShares Barclays 20+ Yr Treas. Bond ETF
        "XLB", # Materials Select Sector SPDR ETF
        "XLE", # Energy Select Sector SPDR ETF
        "XLF", # Financial Select Sector SPDR ETF
        "XLI", # Industrials Select Sector SPDR ETF
        "XLK", # Technology  Select Sector SPDR ETF
        "XLP", # Consumer Staples  Select Sector SPDR ETF
        "XLU", # Utilities  Select Sector SPDR ETF
        "XLV", # Health Care  Select Sector SPDR ETF
        "XLY" # Consumer Discretionary  Select Sector SPDR ETF
  )
}
```
And then global symbols (INTERNATIONAL)
```{r}
global_symbols <- function(){
  symbols <- c(
    enhanced_symbols(),
    "EFA", # iShares EAFE
        "EPP", # iShares Pacific Ex Japan
        "EWA", # iShares Australia
        "EWC", # iShares Canada
        "EWG", # iShares Germany
        "EWH", # iShares Hong Kong
        "EWJ", # iShares Japan
        "EWS", # iShares Singapore
        "EWT", # iShares Taiwan
        "EWU", # iShares UK
        "EWY", # iShares South Korea
        "EWZ", # iShares Brazil
        "EZU", # iShares MSCI EMU ETF
        "IGE", # iShares North American Natural Resources
        "IYR", # iShares U.S. Real Estate
        "IYZ", # iShares U.S. Telecom
        "LQD", # iShares Investment Grade Corporate Bonds
        "SHY" # iShares 42372 year TBonds
  )
}
```
# CheckBLOTTER()
This checks between account object and portfolio object. Credit : Guy Yellin
```{r}
checkBlotterUpdate <- function(port.st = portfolio.st, 
                               account.st = account.st, 
                               verbose = TRUE) {
    
    
  ok <- TRUE
    p <- getPortfolio(port.st)
    a <- getAccount(account.st)
    syms <- names(p$symbols)
    port.tot <- sum(
        sapply(
            syms, 
            FUN = function(x) eval(
                parse(
                    text = paste("sum(p$symbols", 
                                 x, 
                                 "posPL.USD$Net.Trading.PL)", 
                                 sep = "$")))))
    
    port.sum.tot <- sum(p$summary$Net.Trading.PL)
    
    if(!isTRUE(all.equal(port.tot, port.sum.tot))) {
        ok <- FALSE
        if(verbose) print("portfolio P&L doesn't match sum of symbols P&L")
    }
    
    initEq <- as.numeric(first(a$summary$End.Eq))
    endEq <- as.numeric(last(a$summary$End.Eq))
    
    if(!isTRUE(all.equal(port.tot, endEq - initEq)) ) {
        ok <- FALSE
        if(verbose) print("portfolio P&L doesn't match account P&L")
    }
    
    if(sum(duplicated(index(p$summary)))) {
        ok <- FALSE
        if(verbose)print("duplicate timestamps in portfolio summary")
        
    }
    
    if(sum(duplicated(index(a$summary)))) {
        ok <- FALSE
        if(verbose) print("duplicate timestamps in account summary")
    }
    return(ok)
}
```

# Get Symbols 
This is where we access yahoo, google, MYSQL, RData to retrieve symbols using the getSymbols() function found in the quanstrat package
```{r}
#Print the symbols we assigned earlier
print(basic_symbols())

# Assign to symbols object 
symbols <- basic_symbols()
```
Before we use the function, there are some terms here
Symbols: vector of symbols
src: string , our source
index.class: POSIXct, sets class of our object
from: string, first date of data retrieval
to:string, lastdate of data retrieval
adjust: same as before - boolean

# Yahoo !
```{r}
getSymbols(Symbols = symbols, 
           src = "yahoo", 
           index.class = "POSIXct",
           from = start_date, 
           to = end_date, 
           adjust = adjustment)

# verify the stock loaded correctly

head(IWM)
tail(IWM)
summary(IWM)

#clear symbols
rm(list= basic_symbols())
```
#Google 
google no longer works

#FRED
FRED has over 11,000 economic datasets at no cost. We can look at the 10 year treasury rate
```{r}
getSymbols(Symbols = "DGS10",src = "FRED")
# plot it
quantmod::chartSeries(DGS10)
```
# Basic Stategy
SMA = Simple Moving Average - Lagging Indicator because it based on past price data. The longer the time, the greater the lag. Calculated as so:

Sum(Closing Price Past X Days)/ # X Days

I am going to use the Luxor Trading Strategy - SMA(10) & SMA(30)

If the SMA(10) indicator is greater than or equal to the SMA(30) indicator we will submit a stoplimit long order to open and close any short positions that may be open. If the SMA(10) is less than the SMA(30) we will submit a stoplimit short order to open and close any open long positions.


If SMA(10) >= SMA(30):
  BTC short, BTO long
Else if SMA(10) < SMA(30):
  STC long, STO short

Load the Symbols once again (yahoo) and definr the meta data for our symbols
```{r}
getSymbols(Symbols = symbols, 
           src = "yahoo", 
           index.class = "POSIXct",
           from = start_date, 
           to = end_date, 
           adjust = adjustment)
# define the meta data
stock(symbols,
      currency = "USD",
      multiplier = 1)
# assign proper protfolio names
portfolio.st <- "Port.Luxor"
account.st <- "Acct.Luxor"
strategy.st <- "Strat.Luxor"

# remove residuals from previous runs
rm.strat(portfolio.st)
rm.strat(account.st)

# initialize portfolio, account and orders
initPortf(name = portfolio.st,
          symbols = symbols,
          initDate = init_date)
initAcct(name = account.st,
         portfolios = portfolio.st,
         initDate = init_date,
         initEq = init_equity)
initOrders(portfolio = portfolio.st, 
           symbols = symbols,
           initDate = init_date)
strategy(strategy.st, store = T)
```
# Adding indicators

Indicators are a just an average of the previous n prices. SMA(10) is just an average of the last 10 closing prices

TTR library has a bunch of trading functions

add.indicator is a function of quantstrat and adds our indicators to our strategy object. For now we’ll use the following parameters:

strategy: As we stored our strategy name in the strategy.st variable all we need to do is pass that variable. Otherwise we would provide a string. Use variables when this will become redundant as we move along.

name: Indicator function; for this example SMA. We only pass the name of the function as a character string. Parameters for the function are passed into the arguments parameter…

arguments: If we look at ?SMA we see required parameters are x and n with the default n being 10. x is the price object. In our example we are using closing prices.

label: Label of the variable that will be added to our dataset. This must be unique for each indicator we add.

Let’s pause for a moment and examine arguments. Notice we’re passing a series of functions to x. If you wanted to access the Close variable of the IWM dataset you would normally do so by calling IWM$Close or IWM[,4]. Here we’re accessing a mktdata data object

mktdata is a special dataset created for each symbol that will store all of our indicators and signals. When the strategy is ran you will see the mktdata object in your environment. It will only exist for the last symbol the strategy executed.

The add.indicator() function (along with add.signal and add.rules which we’ll discuss momentarily) is not evaluated until we run our strategy. All it does is add our specs to the strategy object. When we run our strategy the mktdata object is created for each symbol iteration where our data will be added.

Cl is actually short-hand for Close as you may have guessed. In fact, we have several short-hand functions for our variables:

Op(): Open

Hi(): High

Lo(): Low

Cl(): Close

Vo(): Volume

Ad(): Adjusted

OpCl(): Open and Close (n x 2 dataset)

HLC(): High, Low and Close (n x 3 dataset)

See the help for any of those symbols above for a more detailed listing.

quote() is a R function that simply wraps the supplied parameter in quotes.

So we’ve added two indicators to our mktdata object, nFast (SMA(10)) and nSlow (SMA(30)). Let’s now add signals.
```{r}
add.indicator(strategy = strategy.st,
              name = "SMA",
              arguments = list(x = quote(Cl(mktdata)),
                               n= 10),
              label = "nFast")
add.indicator(strategy = strategy.st, 
             name = "SMA",
             arguments = list(x = quote(Cl(mktdata)),
                              n = 30),
             label = "nSlow"
             )
```
#Add Signals

Signals are a value given when conditions are met by our indicators. For example, in this strategy we want a signal whenever nFast is greater than or equal to nSlow. We also want another signal where nFast is less than nSlow. We’ll name these signals long and short, respectively.
```{r}
# greater than or equal to signal
add.signal(strategy = strategy.st,
           name = "sigCrossover",
           arguments = list(columns = c("nFast", "nSlow"),
                            relationship = "gte"),
           label = "long")
# less than signal
add.signal(strategy = strategy.st,
           name = "sigCrossover",
           arguments = list(columns = c("nFast", "nSlow"),
                            relationship = "lt"),
           label = "short")
```
Again, we’re passing strategy.st to the strategy parameter. name takes a function just as it did in add.indicator. Here we’ll use some built-in quantstrat functions. Let’s take a quick look at what’s available:

sigComparison: boolean, compare two variables by relationship
gt greater than
lt less than
eq equal to
gte greater than or equal to
lte less than or equal to
sigCrossover: boolean, TRUE when one signal crosses another. Uses the same relationships as sigComparison

sigFormula: apply a formula to multiple variables.

sigPeak: identify local minima or maxima of an indicator

sigThreshold: boolean, when an indicator crosses a value. Uses relationships as identified above.

sigTimestamp: generates a signal based on a timestamp.

We’ll attempt to use each of these signals throughout the book when possible.

#Add Rules

We’ve now constructed our nFast and nSlow indicators and generated signals based on those indicators. Now we have to add rules for those signals.

add.rules will determine the positions we take depending on our signals, what type of order we’ll place and how many shares we will buy.

Whenever our long variable (sigcol) is TRUE (sigval) we want to place a stoplimit order (ordertype). Our preference is at the High (prefer) plus threshold. We want to buy 100 shares (orderqty). A new variable EnterLONG will be added to mktdata. When we enter (type) a position EnterLONG will be TRUE, otherwise FALSE. This order will not replace any other open orders.

```{r}
add.rule(strategy = strategy.st,
         name = "ruleSignal",
         arguments = list(sigcol = "long",
                          sigval = TRUE,
                          orderqty = 100,
                          ordertype = "stoplimit",
                          orderside = "long",
                          threshold = .0005,
                          prefer = "High",
                          TxnFees = -10,
                          replace = FALSE
                          ),
         type = "enter",
         label = "EnterLONG")
```
If our short variable (sigcol) is TRUE (sigval) we will place another stoplimit order (ordertype) with a preference on the Low (prefer). We will sell 100 shares (orderqty). This order will not replace any open orders (replace).

```{r}
add.rule(strategy.st,
         name = "ruleSignal",
         arguments = list(sigcol = "short",
                                 sigval = TRUE,
                                 orderqty = -100,
                                 ordertype = "stoplimit",
                                 threshold = -.005,
                                 orderside = "short",
                                 replace = FALSE,
                                 TxnFees = -10,
                                 prefer = "Low"),
         type = "enter",
         label = "EnterSHORT")
```
Create the rules to leave these open positions

Our next rule, Exit2SHORT, is a simple market order to exit (type) when short is TRUE (sigcol, sigval). This closes out all long positions (orderside, orderqty). This order will replace (replace) any open orders.
```{r}
add.rule(strategy.st, 
         name = "ruleSignal",
         arguments = list(sigcol = "short",
                          sigval = TRUE,
                          orderside = "long",
                          ordertype = "market",
                          orderqty = "all",
                          TxnFees = -10,
                          replace = TRUE),
         type = "exit",
         label = "Exit2SHORT")

```
 Close out any short positions 
 
```{r}
add.rule(strategy.st,
         name = "ruleSignal",
         arguments = list(sigcol = "long",
                          sigval = TRUE,
                          orderside = "short",
                          ordertype = "market",
                          orderqty = "all",
                          TxnFees = -10,
                          replace = TRUE),
         type = "exit",
         label = "Exit2LONG")
```
TxnFees are transaction fees associated with an order. This can be any value you choose but should accurately reflect the fees charged by your selected broker. In addition, we only show them here on exits. Some brokers charge fees on entry positions as well. TxnFees can be added to any rule set.

If you’re not sure what fees your selected broker charges - what’s wrong with you? Go find out now. Some retail brokers (TD Ameritrade, ETrade) will charge under $10 per position on unlimited shares; some such as Interactive Brokers or TradeStation will charge even less depending on the number of shares. $10 is a good starting point.

#Apply Strategy

```{r}
cwd <- getwd()
setwd("./_data")
results_file <- paste("results", strategy.st, "RData", sep = ".")
if( file.exists(results_file) ) {
    load(results_file)
} else {
    results <- applyStrategy(strategy.st, portfolios = portfolio.st)
    updatePortf(portfolio.st)
    updateAcct(account.st)
    updateEndEq(account.st)
    if(checkBlotterUpdate(portfolio.st, account.st, verbose = TRUE)) {
        save(list = "results", file = results_file)
        save.strategy(strategy.st)
    }
}
setwd(cwd)

```
applyStrategy() is the function we will run when we have a straight strategy. What I mean by that is a strategy that doesn’t test different parameters. We’ll get to that type of testing later.

You can see it’s a pretty simple call; we just pass our strategy.st variable as the first parameter and our portfolio as the second parameter. There is no need to get into additional parameters at the moment.

We won’t show the results of any more applyStrategy runs to save space. Just know that if you get trade output you should be good.

Next we update our portfolio and account objects. We do this with the updatePortf(), updateAcct() and updateEndEq() functions. updatePortf calculates the P&L for each symbol in symbols. updateAcct calculcates the equity from the portfolio data. And updateEndEq updates the ending equity for the account. They must be called in order.

We also use the checkBlotterUpdate() mentioned in 3.3. We’re looking for a TRUE value to be returned. Anything FALSE will need to be researched. (If you forgot to clear our your portfolio or strategy with the rm.strat() call mentioned earlier this can result in a FALSE value).

If checkBlotterUpdate returns true we save the results and our strategy (save.strategy) as a RData file into our _data directory. We’ll use them for analysis later.





# DATA QUALITY - Yahoo!
```{r}
getSymbols("SPY",
           src = "yahoo",
           index.class = c("POSIXt", "POSIXct"),
           from = "2010-01-01",
           to = "2011-01-01",
           adjust = TRUE)
yahoo.SPY <- SPY 
summary(yahoo.SPY)
#chart
chart.Posn(portfolio.st, Symbol = "SPY", Dates="2008-01-01::2008-07-01", 
           TA="add_SMA(n = 10, col = 2); add_SMA(n = 30, col = 4)")
```

# Checking the Trades
```{r}
# reload the luxor strategy from before
rm.strat(portfolio.st)
rm.strat(account.st)
symbols <- basic_symbols()
getSymbols(Symbols = symbols, src = "yahoo", index.class = "POSIXct", 
           from = start_date, to = end_date, adjust = adjustment)
initPortf(name = portfolio.st, symbols = symbols, initDate = init_date)
initAcct(name = account.st, portfolios = portfolio.st, initDate = init_date, 
         initEq = init_equity)
initOrders(portfolio = portfolio.st, symbols = symbols, initDate = init_date)
applyStrategy(strategy.st, portfolios = portfolio.st)
checkBlotterUpdate(portfolio.st, account.st, verbose = TRUE)
updatePortf(portfolio.st)
updateAcct(account.st)
updateEndEq(account.st)
```


