---
title: "Algorithmic Trading"
author: "Alexander Martinez"
date: "6/13/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Loading the packages
```{r}
library(quantmod)
library(quantstrat)
library(data.table)
library(DT)
library(ggplot2)
library(htmltools)
library(htmlwidgets)
library(knitr)
library(lattice)
library(pander)
library(tidyverse)
sessionInfo()
```
#Setting system up
```{r}
# Set the time zone to UTC
Sys.setenv(TZ = 'UTC')
#Set currency to USD
currency("USD")
```
#Creating the Objects to Get started
I am going to use a time period that the market is weak. This is because we want to backtest to when times are bad as well as good to get current results.

init_date = date before start date of data retrieval
start_date = date to start data retrieval
end_date = date to end data retrieval 
init_equity = initial account equity
adjustment = TRUE if we should adjust prices for a various amiunt of factors
```{r}
init_date <- "2007-12-13"
start_date <- "2008-01-08"
end_date <- "2009-12-31"
init_equity <- 1e4 #10,000
adjustment <- TRUE
```
# Symbols
Most of the strategies will be focused on 3 ETFs: IWM, QQQ, SPY
```{r}
basic_symbols <- function(){
  symbols <- c("IWM", "QQQ", "SPY")
}
```
To test strategies on a broader scale, we will use enhanced_symbols()
```{r}
enhanced_symbols <- function(){
  symbols <- c(
    basic_symbols(),
    "TLT", # iShares Barclays 20+ Yr Treas. Bond ETF
        "XLB", # Materials Select Sector SPDR ETF
        "XLE", # Energy Select Sector SPDR ETF
        "XLF", # Financial Select Sector SPDR ETF
        "XLI", # Industrials Select Sector SPDR ETF
        "XLK", # Technology  Select Sector SPDR ETF
        "XLP", # Consumer Staples  Select Sector SPDR ETF
        "XLU", # Utilities  Select Sector SPDR ETF
        "XLV", # Health Care  Select Sector SPDR ETF
        "XLY" # Consumer Discretionary  Select Sector SPDR ETF
  )
}
```
And then global symbols (INTERNATIONAL)
```{r}
global_symbols <- function(){
  symbols <- c(
    enhanced_symbols(),
    "EFA", # iShares EAFE
        "EPP", # iShares Pacific Ex Japan
        "EWA", # iShares Australia
        "EWC", # iShares Canada
        "EWG", # iShares Germany
        "EWH", # iShares Hong Kong
        "EWJ", # iShares Japan
        "EWS", # iShares Singapore
        "EWT", # iShares Taiwan
        "EWU", # iShares UK
        "EWY", # iShares South Korea
        "EWZ", # iShares Brazil
        "EZU", # iShares MSCI EMU ETF
        "IGE", # iShares North American Natural Resources
        "IYR", # iShares U.S. Real Estate
        "IYZ", # iShares U.S. Telecom
        "LQD", # iShares Investment Grade Corporate Bonds
        "SHY" # iShares 42372 year TBonds
  )
}
```
# CheckBLOTTER()
This checks between account object and portfolio object. Credit : Guy Yellin
```{r}
checkBlotterUpdate <- function(port.st = portfolio.st, 
                               account.st = account.st, 
                               verbose = TRUE) {
    
    ok <- TRUE
    p <- getPortfolio(port.st)
    a <- getAccount(account.st)
    syms <- names(p$symbols)
    port.tot <- sum(
        sapply(
            syms, 
            FUN = function(x) eval(
                parse(
                    text = paste("sum(p$symbols", 
                                 x, 
                                 "posPL.USD$Net.Trading.PL)", 
                                 sep = "$")))))
    
    port.sum.tot <- sum(p$summary$Net.Trading.PL)
    
    if(!isTRUE(all.equal(port.tot, port.sum.tot))) {
        ok <- FALSE
        if(verbose) print("portfolio P&L doesn't match sum of symbols P&L")
    }
    
    initEq <- as.numeric(first(a$summary$End.Eq))
    endEq <- as.numeric(last(a$summary$End.Eq))
    
    if(!isTRUE(all.equal(port.tot, endEq - initEq)) ) {
        ok <- FALSE
        if(verbose) print("portfolio P&L doesn't match account P&L")
    }
    
    if(sum(duplicated(index(p$summary)))) {
        ok <- FALSE
        if(verbose)print("duplicate timestamps in portfolio summary")
        
    }
    
    if(sum(duplicated(index(a$summary)))) {
        ok <- FALSE
        if(verbose) print("duplicate timestamps in account summary")
    }
    return(ok)
}
```

# Get Symbols 
This is where we access yahoo, google, MYSQL, RData to retrieve symbols using the getSymbols() function found in the quanstrat package
```{r}
#Print the symbols we assigned earlier
print(basic_symbols())

# Assign to symbols object 
symbols <- basic_symbols()
```
Before we use the function, there are some terms here
Symbols: vector of symbols
src: string , our source
index.class: POSIXct, sets class of our object
from: string, first date of data retrieval
to:string, lastdate of data retrieval
adjust: same as before - boolean

# Yahoo !
```{r}
getSymbols(Symbols = symbols, 
           src = "yahoo", 
           index.class = "POSIXct",
           from = start_date, 
           to = end_date, 
           adjust = adjustment)

# get the stock you want
head(IWM)
```



 



