---
title: "Reproducible-Finance-in-R"
author: "Alexander Martinez"
date: "7/27/2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Packages
```{r}
library(PerformanceAnalytics)
library(quantmod)
library(xts)
library(tidyverse)# dplyr, ggplot, tibble, tidyr
library(tidyquant)
library(timetk)# great for converting xts to tibble or tibble to xts
library(tibbletime) # for time aware tibbles and rolling functions

library(highcharter)
```


# Project: Analyze and visualize data for 5 ETFs



Import Data
+ 5 ETFs
 - SPY
 - EFA
 - IJS
 - EEM
 - AGG
 
Import the Prices from Yahoo
```{r}
symbols <- c("SPY","EFA", "IJS", "EEM","AGG")
prices <- getSymbols(symbols, 
                     src = "yahoo",
                     from = "2012-12-31",
                     to = "2017-12-31",
                     auto.assign = TRUE,
                     warnings = FALSE) %>% 
  map(~Ad(get(.))) %>% 
  reduce(merge) %>% 
  `colnames<-`(symbols)
dim(prices)
  
```
### Convert Individual Daily Prices Prices to Portfolio Monthly Return

```{r}
w <- c(.25,
       .25,
       .20,
       .20,
       .10)
```

Portfolio Returns in the 'xts' world

Convert monthly returns

```{r}
prices_monthly <- to.monthly(prices, indexAt = "lastof", OHLC = FALSE) # get the prices of last of every month

asset_returns_xts <- na.omit(Return.calculate(prices_monthly, method = "log"))


```


Convert to portfolio returns
```{r}
portfolio_returns_xts_rebalanced_monthly <-  Return.portfolio(asset_returns_xts,
                                                              weights = w,
                                                              rebalance_on = "months") %>%
  `colnames<-`("returns")

portfolio_returns_xts_rebalanced_monthly %>% 
  head()
# dates in xts structure are stored in index
index(portfolio_returns_xts_rebalanced_monthly) 


```




Portfolio returns in the tidyverse

First get monthly returns of individual ETFs
```{r}
asset_returns_dplyr_byhand <- 
  prices %>%
  to.monthly(indexAt = "lastof", OHLC = F) %>%
  #convert index to date 
  data.frame(date = index(.)) %>%
  remove_rownames() %>%
  gather(asset, returns, -date) %>%
  group_by(asset) %>%
  mutate(returns = (log(returns) - lag(log(returns))))%>%
  # this spreads the each asset class into a column and then it 
  # puts the returns on each asset in that column
  spread(asset,returns) %>%
  select(date, symbols) %>%
  na.omit()
```
Use `case_when()` to assign weights
```{r}
portfolio_returns_dplyr_byhand <-
  asset_returns_dplyr_byhand %>%
  gather(asset, returns, -date)%>%
  group_by(asset) %>%
  mutate(weights = case_when(asset == symbols[1] ~ w[1],
                             asset == symbols[2] ~ w[2],
                             asset == symbols[3] ~ w[3],
                             asset == symbols[4] ~ w[4],
                             asset == symbols[5] ~ w[5]),
         weighted_returns = returns * weights) %>%
  group_by(date) %>%
  summarise(returns = sum(weighted_returns))

portfolio_returns_dplyr_byhand %>% 
  head()
```

Portfolio returns in the tidyquant world using `tq_portfolio()`

```{r}
portfolio_returns_tq_rebalanced_monthly <- 
  asset_returns_dplyr_byhand %>%
  gather(asset, returns, -date) %>%
  tq_portfolio(assets_col = asset,
               returns_col = returns,
               weights = w,
               col_rename = "returns",
               rebalance_on = "months")
portfolio_returns_tq_rebalanced_monthly %>%
  head()
```
Always check for consistent results ie make sure the returns are all the same


### Portfolio Risk, or Vol, or Standard Deviation

Calculate Risk, volatility, of portfolio

We will go through how to do so using `xts`, tidyverse and `tidyquant`,

Standard Deviation in the `xts` world

use built in stdev function

```{r}
portfolio_sd_xts_builtin <- 
  StdDev(portfolio_returns_xts_rebalanced_monthly)

portfolio_sd_xts_builtin_percent <- 
  round(portfolio_sd_xts_builtin * 100, 2) 

portfolio_sd_xts_builtin_percent
```

other builtin functions that would be extremely useful
```{r}
table.DownsideRisk(portfolio_returns_xts_rebalanced_monthly)
table.Drawdowns(portfolio_returns_xts_rebalanced_monthly)
SemiDeviation(portfolio_returns_xts_rebalanced_monthly)
SharpeRatio(portfolio_returns_xts_rebalanced_monthly)
```


```{r}
portfolio_sd_tidy_builtin_percent <- 
  portfolio_returns_dplyr_byhand %>%
  summarise(
    sd = sd(returns),
    sd_byhand = sqrt(sum((returns - mean(returns))^2)/(nrow(.)-1))) %>%
  mutate(dplyr = round(sd, 4)* 100,
         dplyr_byhand = round(sd_byhand, 4) * 100)

portfolio_sd_tidy_builtin_percent
```

`Tidyquant` is a good bridge between these two using `table.Stats`
```{r}
portfolio_sd_tidyquant_builtin_percent <-
  portfolio_returns_tq_rebalanced_monthly %>%
  tq_performance(Ra = returns,
                 Rb = NULL,
                 performance_fun = table.Stats)%>%
  select(Stdev) %>%
  mutate(tq_sd = round(Stdev, 4) * 100)

portfolio_sd_tidyquant_builtin_percent
```
Visualizing Standard Deviation and Returns

Returns

```{r}
portfolio_returns_dplyr_byhand %>%
  ggplot(aes(x = date, y = returns)) +
  geom_point(color = "cornflowerblue") +
  scale_x_date(breaks = scales::pretty_breaks(n = 6)) +
  ggtitle("Scatterplot of Returns by Date") +
  theme(plot.title = element_text(hjust = .5))
```
SD plot
```{r}
sd_plot <-
  sd(portfolio_returns_tq_rebalanced_monthly$returns)

mean_plot <- 
  mean(portfolio_returns_tq_rebalanced_monthly$returns)

portfolio_returns_tq_rebalanced_monthly %>%
  mutate(hist_col_red = ifelse(returns < (mean_plot - sd_plot),
                               returns, as.numeric(NA)),
         hist_col_green = ifelse(returns > (mean_plot + sd_plot),
                               returns, as.numeric(NA)),
         hist_col_blue = ifelse(returns > (mean_plot - sd_plot) & 
                                  returns < (mean_plot +  sd_plot), returns, as.numeric(NA)))%>%
  ggplot(aes(x = date)) +
  geom_point(aes(y = hist_col_red),
             color = "red") +
  geom_point(aes(y = hist_col_blue),
             color = "blue") +
  geom_point(aes(y = hist_col_green),
             color = "green") +
  labs(title = "Colored Scatter", y = "monthly returns") +
  scale_color_date(breaks = scales::pretty_breaks(n = 8))+
  theme(plot.title = element_text(hjust = .5))
  
```
```{r}
asset_returns_dplyr_byhand %>%
  gather(asset, returns, -date) %>%
  group_by(asset) %>%
  summarize(sd = 100 * sd(returns)) %>%
  add_row(asset = "Portfolio",
          sd = portfolio_sd_tidy_builtin_percent$dplyr) %>%
  ggplot(aes(x= asset,
             y = sd,
             color = asset)) +
  geom_point() +
  geom_text(aes(label = asset),
            nudge_y = .2) +
  labs(title = "Std Dev Returns of Each Asset",
       y = "standard dev") +
  scale_y_continuous(labels = function(x) paste0(x, "%"))+
  theme(plot.title = element_text(hjust = .5))


# the end
```
 

